---
title: "AE 03: Duke Forest + data visualization"
author:"Isabel Ulrich"
editor: 
  markdown: 
    wrap: sentence
---

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(openintro)
library(dplyr)
library(stringr)
```

## Exercise 1

Suppose you're helping some family friends who are looking to buy a house in Duke Forest. As they browse Zillow listings, they realize some houses have garages and others don't, and they wonder: **Does having a garage make a difference?**

Luckily, you can help them answer this question with data visualization!

-   Make histograms of the prices of houses in Duke Forest based on whether they have a garage.
    -   In order to do this, you will first need to create a new variable called `garage` (with levels `"Garage"` and `"No garage"`).
    -   Below is the code for creating this new variable. Here, we `mutate()` the `duke_forest` data frame to add a new variable called `garage` which takes the value `"Garage"` if the text string `"Garage"` is detected in the `parking` variable and takes the test string `"No garage"` if not.

```{r}
#| label: garage
#| eval: false

duke_forest <- duke_forest |>
  mutate(garage = if_else(str_detect(parking, "Garage"),   "Garage", "No garage"))
```

-   Then, facet by `garage` and use different colors for the two facets.
-   Choose an appropriate binwidth and decide whether a legend is needed, and turn it off if not.
-   Include informative title and axis labels.
-   Finally, include a brief (2-3 sentence) narrative comparing the distributions of prices of Duke Forest houses that do and don't have garages. Your narrative should touch on whether having a garage "makes a difference" in terms of the price of the house.

```{r}
#| label: price-garage

ggplot(duke_forest, aes(x=price, fill=garage))+
  geom_histogram(binwidth=50000)+
  facet_wrap(~garage)+
  labs(x="Price of House ($)", title = "Prices of Houses in Duke Forest with Garages and no Garages")+
  theme(legend.position = "none")
```

Based on the graph there seems to be no large differnce in prices of house on weather it has a garage or not. Both graphs show an average price around \$500000. The price of houses does not change between having a garage or not.

::: callout-important
Now is a good time to render, commit, and push. Make sure that you commit and push all changed documents and your Git pane is completely empty before proceeding.
:::

## Exercise 2

It's expected that within any given market larger houses will be priced higher. It's also expected that the age of the house will have an effect on the price. However in some markets new houses might be more expensive while in others new construction might mean "no character" and hence be less expensive. So your family friends ask: "In Duke Forest, do houses that are bigger and more expensive tend to be newer ones than those that are smaller and cheaper?"

Once again, data visualization skills to the rescue!

-   Create a scatter plot to exploring the relationship between `price` and `area`, conditioning for `year_built`.
-   Use `geom_smooth()` with the argument `se = FALSE` to add a smooth curve fit to the data and color the points by `year_built`.
-   Include informative title, axis, and legend labels.
-   Discuss each of the following claims (1-2 sentences per claim). Your discussion should touch on specific things you observe in your plot as evidence for or against the claims.
    -   Claim 1: Larger houses are priced higher.
    -   Claim 2: Newer houses are priced higher.
    -   Claim 3: Bigger and more expensive houses tend to be newer ones than smaller and cheaper ones.

```{r}
#| label: price-area

ggplot(duke_forest, aes(x=price, y=area,color=year_built))+
geom_point()+
geom_smooth(se=FALSE)+
labs(x="Price of House ($)",y="Area of House"
     ,title="Price vs Area of Houses in Duke Forest by Age", 
     color="Year House Built")
```

Claim 1: Larger houses are priced higher:

-   Following the smooth curve fit to the data, there is a positive relation to area and price. The trend is that as the price increases the area a house increases as well. It could be concluded that larger houses are priced higher; however there are outliers that don't match this claim. Such as the house "2832 Mcdowell Rd, Durham, NC 27705" is on the lower end of price at \~\$10000 but has similar area ( \~5000) of houses at \~\$1200000. It is not always the case that a larger house is a higher price.

Claim 2: Newer houses are priced higher.

-   Based on the scatterplot at the extremes begining of the data's range the darker plots (houses built longer ago/older houses) are at lower prices. However the newest (light blue plots) can be found at a lower price then other darker plots. The claim that newer houses are priced higher is not a strong claim as there are many cases shown that older houses are prices higher then the newer houses.

Claim 3: Bigger and more expensive houses tend to be newer ones than smaller and cheaper ones.

-   Based on the scatterplot a trend can be seen that the houses at the very low ranger (smallest and cheapest) tend to be darker plots (older houses) and as the price and area of the house increases there are instances where the houses are newer but not always. As the area and price increases it is a mix of older and newer houses. Thus claiming that bigger and more expensive houses tend to be newer ones then smaller cheaper ones is not a strong claim.

::: callout-important
Now is a good time to render, commit, and push. Make sure that you commit and push all changed documents and your Git pane is completely empty before proceeding.
:::
